Retrofitting Effect Handlers onto OCaml

### 2020-11-13 retro-concurrency.tex
Fix typo in code
Update abstract

\begin{abstract}
Effect handlers have been gathering momentum as a mechanism
for modular programming with user-defined effects. The runtime semantics of
effect handlers allow for control flow
mechanisms such as first-class continuations, generators, async/await and coroutines to be composably expressed. We present a design and evaluate a
full-fledged efficient implementation of effect handlers for OCaml, an industrial-strength multi-paradigm programming language. Our implementation strives to maintain the existing performance profile of conventional OCaml code, particularly in the domain of systems and toolchain programming. Retrofitting effect handlers onto OCaml is challenging since OCaml does currently not have any non-local control flow mechanisms other than exceptions. Our implementation of
effect handlers for OCaml: {\em (i)} imposes negligible overhead on code that does not use
effect handlers; {\em (ii)} remains compatible with program analysis tools that inspect the
stack; and {\em (iii)} is efficient for new code that makes use of effect handlers.
\end{abstract}

@@

We consider well-behaved Multicore OCaml programs to resume their continuations
\emph{exactly once}. This can be enforced by installing a finaliser on every
captured continuation that |discontinue|s the continuation during finalisation
and ignores the result:
\begin{lstlisting}
Gc.finalise (fun k ->
  try ignore (discontinue k Unwind) | _ -> ()) k
\end{lstlisting}
Multicore OCaml does not enforce this behaviour in order to avoid installing an
effect handler every time a continuation is captured.

>>>>

\begin{abstract}
Effect handlers have been gathering momentum as a mechanism for modular
	programming with user-defined effects. Effect handlers allow for non-local
	control flow mechanisms such as generators, async/await, lightweight threads
	and coroutines to be composably expressed. We present a design and evaluate a
	full-fledged efficient implementation of effect handlers for OCaml, an
	industrial-strength multi-paradigm programming language. Our implementation
	strives to maintain the backwards compatibility and existing performance
	profile of conventional OCaml code. Retrofitting effect handlers onto OCaml
	is challenging since OCaml does currently not have any non-local control flow
	mechanisms other than exceptions. Our implementation of effect handlers for
	OCaml: {\em (i)} imposes negligible overhead on code that does not use effect
	handlers; {\em (ii)} remains compatible with program analysis tools that
	inspect the stack; and {\em (iii)} is efficient for new code that makes use
	of effect handlers.
\end{abstract}

@@

We consider well-behaved Multicore OCaml programs to resume their continuations
\emph{exactly once}. This can be enforced by installing a finaliser on every
captured continuation that |discontinue|s the continuation during finalisation
and ignores the result:
\begin{lstlisting}
Gc.finalise (fun k ->
  try ignore (discontinue k Unwind) with _ -> ()) k
\end{lstlisting}
Multicore OCaml does not enforce this behaviour in order to avoid installing an
effect handler every time a continuation is captured.
